Industrial, large-scale model predictive control with structured neural networks

### 2020-11-03 paper.tex
fix norm, xhat, dhat, and steady state adjectives.

Notation. The symbols $\bbI$ and $\bbR$ are 
used to denote integers and reals respectively. 
Subscripts denote restrictions 
(e.g., $\bbR_{\geq 0}$ for nonnegative reals and 
$\bbI_{a:b}$ for integers in the 
closed interval $[a, b]$). The Euclidean norm 
is denoted by $\vert \cdot \vert$. A function 
$\alpha : \bbR_{\geq 0} \rightarrow \bbR_{\geq 0}$
is of class $\mathcal{K}$ if it is continuous, zero 
at zero, and strictly increasing. This function is of class 
$\mathcal{K}_{\infty}$ if it is of class $\mathcal{K}$
and unbounded ($\alpha(s) \rightarrow \infty$ as 
$s \rightarrow \infty$). A function
$\beta : \bbR_{\geq 0} \times \bbI_{\geq 0} \rightarrow \bbR_{\geq 0}$
is of class $\mathcal{KL}$ if it is continuous,
for each $k \in \bbI_{\geq 0}$, $\beta(\cdot, k)$ is of class $\mathcal{K}$,
and for each $s \geq 0$, $\beta(s, \cdot)$ is nonincreasing
and satisfies $\textnormal{lim}_{k \rightarrow \infty} \beta(s, k) = 0$.
Given $V : X \rightarrow \bbR_{\geq 0}$ and $\tau > 0$, define 
$\textnormal{lev}_{\tau} V = \{ x \in X \mid V(x) \leq \tau \}$.
Bold symbols, e.g., $\mathbf{d}$ denote sequences, 
$d(k)$ denotes an element of $\mathbf{d}$ at time 
$k \in \bbI_{\geq 0}$, and $\mathbf{d}_i$ 
denotes the collection of elements of 
$\mathbf{d}$ for $k \in \bbI_{0:i-1}$.
Define $\vert \vert \mathbf{d}_i \vert \vert = 
\textnormal{max}_{k \in \bbI_{0:i-1}} \vert d(k) \vert$.
For a given vector $a$, lower and upper bounds 
($\underline{a}, \overline{a}$), define the saturation function 
as $\textnormal{sat}(a, \underline{a}, \overline{a}) = 
\{ a \ \textnormal{if} \ \underline{a} \leq a \leq \overline{a}; 
\underline{a} \ \textnormal{if} \ a < \underline{a};
\overline{a} \ \textnormal{if} \ \overline{a} < a \}$.

@@

\subsection{Model}
We consider a linear discrete time model augmented with an 
integrating disturbance model \citep{pannocchia:rawlings:2003}
\begin{align}\label{eq:ltimodel}
    x^+ = Ax + Bu + B_dd, \quad d^+ = d, \quad y = Cx + C_dd
\end{align}
in which $x \in \bbR^n$ is the state, 
$u \in \bbR^m$ is the control input, and
$d \in \bbR^d$ is the state of the disturbance
model. The matrices $A \in \bbR^{n \times n}$, 
$B \in \bbR^{n \times m}$,  
$C \in \bbR^{p \times n}$ denote the actuator to measurement 
model, and $B_d \in \bbR^{n \times d}$, $C_d \in \bbR^{p \times d}$
is the disturbance model.
The objective of the disturbance model
is to remove offset in the controlled 
plant measurements at steady
state operations and maintain
the controlled variables at their setpoints.
We assume in the rest of this 
article that a Kalman filter can be constructed to 
estimate the model state ($\hat{x}$) and disturbance ($\hat{d}$) 
from measurements.

\subsection{Target selector}
Given the disturbance estimate, the
input and controlled measurement setpoints,
we consider the following target selector QP
\begin{align}
    \min_{x_s, \ u_s}  \dfrac{1}{2} |u_{sp} - u_s|^2_{R_s}
\end{align}
subject to
\begin{align}  
    \begin{bmatrix}
        I -A & -B \\
        HC & 0
    \end{bmatrix} \begin{bmatrix}
        x_s \\
        u_s
    \end{bmatrix} &= \begin{bmatrix}
        B_d\hat{d} \\
        r_{sp} - HC_d\hat{d}
    \end{bmatrix} \\
    \underline{u} \leq u_s \leq \overline{u}
\end{align}
in which
($x_s, u_s$) is the 
target steady state pair,
$r=Hy$ are the controlled measurements
chosen as some subset 
or a linear combination of all the measurements,
($\underline{u}, \overline{u}$) are the 
actuator bounds, and
($u_{sp}, r_{sp}$) are the input and 
controlled measurement setpoints.
The equality constraint $r_{sp} = H(Cx_s + C_d\hat{d})$
can be difficult to satisfy exactly in real time 
depending on the controlled measurement setpoint 
and disturbance estimate. 
Then, the hard setpoint equality constraint 
can be relaxed and moved to the stage cost such 
that the target selector computes a steady state
to minimize the offset in the controlled measurements. 
We assume for the simulation studies
in this article that the input setpoint ($u_{sp}$)
is fixed at some chosen steady state and 
only the controlled measurement setpoint ($r_{sp}$) 
changes in real time.

\subsection{Regulator}
Based on the state estimate and
target steady state pair,
the following QP is solved to determine the 
actuator move to be applied to the plant
\begin{align} \label{eqn:regulator}
    V_N(\tilde{x}(0), \mathbf{\tilde{u}}) = \dfrac{1}{2}\sum_{k=0}^{N-1} & \Big(|\tilde{x}(k)|^2_Q  + |\tilde{u}(k)|^2_R \Big) + \dfrac{1}{2}|\tilde{x}(N)|^2_P
\end{align}
subject to 
\begin{align}
    \tilde{x}^+ &= A\tilde{x} + B\tilde{u}, \quad \tilde{x}(0) = \hat{x} - x_s \\
    &\underline{u} \leq \tilde{u} + u_s \leq \overline{u}
    \label{eqn:regulator_constraints}
\end{align}
in which $\hat{x}$ is the state estimate after the current measurement,
$N$ is the control 
horizon length, and
($\tilde{x}, \tilde{u}$) are the state and control 
in deviation from the current steady state targets ($x_s, u_s$). 
The penalty matrix $P$ is chosen as the 
optimal cost-to-go matrix of the unconstrained infinite horizon 
linear quadratic problem.
The decision variable in this 
QP is the control sequence 
$\mathbf{\tilde{u}} = [\tilde{u}(0)', \tilde{u}(1)', ..., \tilde{u}(N-1)']'$.
The first move of the optimal solution is denoted 
as the MPC feedback law,
$\kappa_N(\hat{x}, x_s, u_s) = 
\mathbf{\tilde{u}}^0(0;\hat{x}, x_s, u_s) + u_s$, 
which is applied to the plant. This feedback law is a function of the 
state estimate and target steady-state pair. A rate-of-change 
penalty on the control can be imposed in the QP objective
by augmenting the dynamic model with a surrogate state 
for the control as discussed in \cite*{rao:rawlings:1999}. 
In this case, 
the control injected at the previous timestep
($u_{-1}$) is supplied to the QP, 
and the feedback law becomes a function of the previous control 
as well in addition to the state estimate
and target steady-state pair.

@@

As reviewed in \cite*{mayne:rawlings:rao:scokaert:2000},
the nominal stability of MPC can be established
using the optimal cost function ($V_N^0(x)$) of the QP
as a Lyapunov function.
The region of attraction of the closed-loop system 
$x^+ = Ax + B\kappa_N(x)$ when no hard terminal region 
constraint is included in the QP
can be characterized as $\mathcal{X}_N = \textnormal{lev}_{Nd + \tau} \
V_N^0(x)$ \citep*{limon:alamo:salas:camacho:2006}.
The parameter $\tau$ is the level set
parameter of the terminal region chosen as
$\bbX_f = \textnormal{lev}_{\tau} \ x'Px$, and $d > 0$
is a constant such that $\ell(x, u) \geq d$ for all 
$x \in \bbR^n \setminus \bbX_f$ and 
$\underline{u} \leq u+u_s \leq \overline{u}$.
The parameter $\tau$ is chosen small enough such that
the unconstrained LQR satisfies the actuator constraints 
for all $x \in \bbX_f$. The optimal cost 
function is continuous and satisfies 
$c_1\vert x \vert^2 \leq V_N^0(x) \leq c_2\vert x \vert^2$
and $V_N^0(x^+) - V_N^0(x) \leq -c_1\vert x \vert^2$ 
for some $c_2 \geq c_1 > 0$.

Assume that a state estimate ($\hat{x}$) is
used by the NN to compute the control. 
To analyze robustness, the following perturbed
linear system is considered
\begin{align} \label{eq:perturbed_sys}
    \hat{x}^+ = A\hat{x} + B\kappa_N(\hat{x}) + Be_{NN}(\hat{x}) + w - Ae + e^+
\end{align}
in which $\hat{x} = x + e$ is the state estimate, 
$w$ is a process disturbance, 
and $e$, $e^+$ are the state estimation errors at the 
current and next timestep respectively. 
We use $\phi_d(k;\hat{x})$ to denote a solution of the perturbed system  at time
$k$ starting from an initial state $\hat{x}$; and $\phi(k;x)$ to denote a
solution  of the nominal system starting  from an initial state $x$ when the
state estimation errors and process disturbances are zero.

\begin{thm} \label{thm:nnrobustness}
For all $ 0 < \rho \leq Nd + \tau$, 
there exist constants $\delta_1, \delta_2, \delta_3 > 0$, 
functions $\beta(\cdot) \in \mathcal{K} \mathcal{L}$
and $\alpha_e(\cdot) , \alpha_w(\cdot), \alpha_{n}(\cdot) \in \mathcal{K}$,
such that for all disturbance sequences
satisfying $\vert\vert \mathbf{e}_{k+1} \vert\vert \leq \delta_1$,
$\vert\vert \mathbf{w}_k \vert\vert \leq \delta_2$,
and $\vert e_{NN}(\hat{x}) \vert \leq \delta_3$,
for all $\hat{x}$ in the set $S := \textnormal{lev}_{\rho} \ V_N^0(\hat{x})$,
we have the following bound 
$\vert \phi_d(k; \hat{x}) \vert \leq \beta(\vert \hat{x} \vert, k) 
+ \alpha_e(\vert \vert\mathbf{e}_{k+1} \vert \vert) + 
\alpha_w(\vert \vert \mathbf{w}_k \vert \vert) + 
\alpha_n(\bar{e}_{NN})$, 
in which 
$\bar{e}_{NN} = \textnormal{max}_{\hat{x} \in S} \vert e_{NN}(\hat{x}) \vert$.
\end{thm}

The proof of this theorem along with the required 
input-to-state stability definitions are provided 
in Appendix \ref{app:theorem1}.
In the nominal case, 
the disturbance sequences $\mathbf{e}$ and $\mathbf{w}$ are zero.
The closed-loop states in this case have the bound
$\vert \phi(k; x) \vert \leq \beta(\vert x \vert, k) + 
\alpha_n(\bar{e}_{NN})$. Any additional value of 
the NN training is to reduce the worst 
approximation error ($\bar{e}_{NN}$) 
in the state space of interest and 
reduce the bound on the closed-loop states.

@@

\begin{itemize}
    \item Controller performance index
    \begin{align*}        
    \Lambda_k = \dfrac{1}{k}\sum_{t=1}^{k} & \big(\vert x(t) - x_{s} (t) \vert^2_{Q} \\
        & + \vert u(t) - u_{s} (t) \vert^2_{R} + \vert \Delta u(t) \vert^2_{S}\big)         
    \end{align*}
    \item \% Performance loss = $100(\Lambda^{\textnormal{F}}_{N_{t}} - \Lambda^{\textnormal{MPC}}_{N_{t}})/\Lambda^{\textnormal{MPC}}_{N_{t}}$
    \item Average and worst-case speedups
    \item Data generation and NN training times
    \item Memory required to store the weights of the NNs
\end{itemize}
Here, $N_t$ is the number of simulation time steps,  
$\Lambda^{\textnormal{F}}_{N_{t}}$ and 
$\Lambda^{\textnormal{MPC}}_{N_{t}}$
are the average stage costs obtained 
at the end of the simulation period by the 
fast controllers and the MPC controller.
The data generation and online timing comparisons are 
performed on a computing cluster that has 
several multi-core CPUs of clock speed 2.4GHz. 
The NN trainings are performed on a 
Tesla V100-SXM2 GPU that has 32 GB memory.

@@

\begin{definition} \label{def:iss}
(Input-to-state stability (ISS)). $\bbW$ is a compact 
set containing the origin and that $X$ is a closed robustly
positive invariant set for the 
system $x^+ = f(x, w)$, $w \in \bbW$. This
system is ISS in $X$ if there exist functions 
$\beta(\cdot) \in \mathcal{K}\mathcal{L}$ and 
$\sigma(\cdot) \in \mathcal{K}$, such that for all
$x \in X$ and $w(i) \in \bbW$, $i \in \bbI_{0:k-1}$, 
$\vert \psi(k; x, \mathbf{w}_k) \vert \leq \beta(\vert x\vert, k) 
+ \sigma(\vert \vert \mathbf{w}_k\vert \vert)$, in which 
$\psi(k; x, \mathbf{w}_k)$ is the solution of the system 
at time $k$, starting from an initial state $x$, and with the
disturbance sequence $\mathbf{w}_k$ affecting the system.
\end{definition}

\begin{definition} \label{def:iss_lyapunov_func}
(ISS-Lyapunov function). A function $V : X \rightarrow \bbR_{\geq 0}$
is an ISS-Lyapunov function in $X$ for the system $x^+ = f(x, w)$, $w \in \bbW$
if there exist functions $\alpha_1(\cdot), \alpha_2(\cdot), \alpha_3(\cdot) 
\in \mathcal{K}_{\infty}$ and $\sigma(\cdot) \in \mathcal{K}$
such that for all $x \in X$ and $w \in \bbW$
\begin{align*}
    \alpha_1(\vert x \vert) \leq & V(x) \leq \alpha_2(\vert x \vert) \\
    V(f(x, w)) &- V(x) \leq -\alpha_3(\vert x \vert) + \sigma(\vert w \vert)
\end{align*}
\end{definition}

@@

\begin{prop} \label{prop:continuous_funcs}
    Let $C \subseteq D \subseteq \bbR^n$, $C$ compact, 
    $D$ closed, and $f : D \rightarrow \bbR^n$ continuous.  
    Then there exists a $\sigma(\cdot) \in \mathcal{K}_{\infty}$,
    such that for all $x \in C$ and $y \in D$, we have that 
    $\vert f(x) - f(y) \vert \leq \sigma(\vert x-y \vert)$.
\end{prop}

@@

\textbf{Proof of Theorem \ref{thm:nnrobustness}:}
We proceed in two steps. First, 
we show that there exist constants $\delta_1, \delta_2, \delta_3 > 0$
such that the set $S$ is robustly positive invariant. Then, we 
show that $V_N^0(\hat{x})$ is an ISS Lyapunov function for 
the perturbed system \eqref{eq:perturbed_sys} on this set $S$. Denote 
$\tilde{x}^+ = A\hat{x} + B\kappa_N(\hat{x})$ as the nominal 
state evolution using the optimal MPC controller, 
and $d(k) = [ e_{NN}(\hat{x}(k))'B', w(k)', e(k)', e(k+1)']'$
as the combination of all the disturbances at a given time $k$.

To show robust positive invariance of $S$,  
we establish that if $V_N^0(\hat{x}) \leq \rho$, 
then $V_N^0(\hat{x}^+) \leq \rho$.
Since the optimal 
cost function ($V_N^0(\hat{x})$) is continuous,
using $\bbR^n$ as the closed set ($\hat{x}^+ \in \bbR^n$) 
and $S$ as the compact set ($\tilde{x} \in S$),
from proposition \ref{prop:continuous_funcs} we have
for some $\sigma_1(\cdot) \in \mathcal{K}_{\infty}$
\begin{align}
    \vert V_N^0(\hat{x}^+) - V_N^0(\tilde{x}^+) \vert &\leq 
    \sigma_1(\vert \hat{x}^+ - \tilde{x}^+ \vert)  \nonumber  \\
    &\leq 
   \sigma_1(\vert B e_{NN}(\hat{x}) \vert + \vert w \vert 
   \nonumber \\
   & + \vert A\vert\vert e \vert +  \vert e^+ \vert) \nonumber \\
     &\leq 
   \sigma_1(\vert d \vert + \vert d \vert + \vert A \vert\vert d \vert + \vert d \vert) := \sigma_2(\vert d \vert)
   \label{eq:optimal_cost_bound}
\end{align}
in which $\sigma_2(s) := \sigma_1(\vert A\vert s +3s)$, and the last inequality 
follows because $\vert a \vert \leq \vert [a', b']' \vert$, in which $a$ and $b$
are both vectors, and $\sigma_2(\cdot) \in \mathcal{K}_{\infty}$.
We partition $S$ in outer and inner parts based on 
the value of the optimal cost $V_N^0(\hat{x})$, and bound the 
combination of disturbances $d$ in each part 
such that $V_N^0(\hat{x}^+) \leq \rho$. \\
Case 1: $\rho/2 \leq V_N^0(\hat{x}) \leq \rho $. Using the 
upper bound on the optimal cost, we have 
$\rho/(2c_2) \leq \vert \hat{x} \vert^2$ for all $\hat{x}$ 
in this part of $S$. To analyze the worst-case disturbance, 
we have from the equation \eqref{eq:optimal_cost_bound} that 
$V_N^0(\hat{x}^+) \leq V_N^0(\tilde{x}^+) + \sigma_2(\vert d \vert) $.
From nominal MPC cost decrease, we obtain  
$V_N^0(\hat{x}^+) \leq V_N^0(\hat{x}) - c_1\vert \hat{x} \vert^2 + \sigma_2(\vert d \vert)$. Thus, if $\sigma_2(\vert d \vert) \leq \rho c_1/(2c_2)$, 
at all times in this outer part of $S$, we have 
$V_N^0(\hat{x}^+) \leq \rho$. \\
Case 2: $V_N^0(\hat{x}) \leq \rho/2 $. From 
equation \eqref{eq:optimal_cost_bound} and nominal MPC 
cost decrease, we have for this part of $S$ that 
$V_N^0(\hat{x}^+) \leq \rho/2 + \sigma_2(\vert d \vert)$.
Thus, if $\sigma_2(\vert d \vert) \leq \rho/2$, 
at all times in this inner part of $S$, we have 
$V_N^0(\hat{x}^+) \leq \rho$.

>>>>

Notation. The symbols $\bbI$ and $\bbR$ are 
used to denote integers and reals respectively. 
Subscripts denote restrictions 
(e.g., $\bbR_{\geq 0}$ for nonnegative reals and 
$\bbI_{a:b}$ for integers in the 
closed interval $[a, b]$). The Euclidean norm 
is denoted by $\norm{\cdot}$. A function 
$\alpha : \bbR_{\geq 0} \rightarrow \bbR_{\geq 0}$
is of class $\mathcal{K}$ if it is continuous, zero 
at zero, and strictly increasing. This function is of class 
$\mathcal{K}_{\infty}$ if it is of class $\mathcal{K}$
and unbounded ($\alpha(s) \rightarrow \infty$ as 
$s \rightarrow \infty$). A function
$\beta : \bbR_{\geq 0} \times \bbI_{\geq 0} \rightarrow \bbR_{\geq 0}$
is of class $\mathcal{KL}$ if it is continuous,
for each $k \in \bbI_{\geq 0}$, $\beta(\cdot, k)$ is of class $\mathcal{K}$,
and for each $s \geq 0$, $\beta(s, \cdot)$ is nonincreasing
and satisfies $\textnormal{lim}_{k \rightarrow \infty} \beta(s, k) = 0$.
Given $V : X \rightarrow \bbR_{\geq 0}$ and $\tau > 0$, define 
$\textnormal{lev}_{\tau} V = \{ x \in X \mid V(x) \leq \tau \}$.
Bold symbols, e.g., $\mathbf{d}$ denote sequences, 
$d(k)$ denotes an element of $\mathbf{d}$ at time 
$k \in \bbI_{\geq 0}$, and $\mathbf{d}_i$ 
denotes the collection of elements of 
$\mathbf{d}$ for $k \in \bbI_{0:i-1}$.
Define $\norm{\norm{\mathbf{d}_i}} = 
\textnormal{max}_{k \in \bbI_{0:i-1}} \norm{d(k)}$.
For a given vector $a$, lower and upper bounds 
($\underline{a}, \overline{a}$), define the saturation function 
as $\textnormal{sat}(a, \underline{a}, \overline{a}) = 
\{ a \ \textnormal{if} \ \underline{a} \leq a \leq \overline{a}; 
\underline{a} \ \textnormal{if} \ a < \underline{a};
\overline{a} \ \textnormal{if} \ \overline{a} < a \}$.

@@

\subsection{Model}
We consider a linear discrete time model augmented with an 
integrating disturbance model \citep{pannocchia:rawlings:2003}
\begin{align}\label{eq:ltimodel}
    x^+ = Ax + Bu + B_dd, \quad d^+ = d, \quad y = Cx + C_dd
\end{align}
in which $x \in \bbR^n$ is the state, 
$u \in \bbR^m$ is the control input, and
$d \in \bbR^d$ is the state of the disturbance
model. The matrices $A \in \bbR^{n \times n}$, 
$B \in \bbR^{n \times m}$,  
$C \in \bbR^{p \times n}$ denote the actuator to measurement 
model, and $B_d \in \bbR^{n \times d}$, $C_d \in \bbR^{p \times d}$
is the disturbance model.
The objective of the disturbance model
is to remove offset in the controlled 
plant measurements at steady
state operations and maintain
the controlled variables at their setpoints.
We assume in the rest of this 
article that a Kalman filter can be constructed to 
estimate the model state ($\xhat$) and disturbance ($\dhat$) 
from measurements.

\subsection{Target selector}
Given the disturbance estimate, the
input and controlled measurement setpoints,
we consider the following target selector QP
\begin{align}
    \min_{x_s, \ u_s}  \dfrac{1}{2} |u_{sp} - u_s|^2_{R_s}
\end{align}
subject to
\begin{align}  
    \begin{bmatrix}
        I -A & -B \\
        HC & 0
    \end{bmatrix} \begin{bmatrix}
        x_s \\
        u_s
    \end{bmatrix} &= \begin{bmatrix}
        B_d\dhat \\
        r_{sp} - HC_d\dhat
    \end{bmatrix} \\
    \underline{u} \leq u_s \leq \overline{u}
\end{align}
in which
($x_s, u_s$) is the 
target steady-state pair,
$r=Hy$ are the controlled measurements
chosen as some subset 
or a linear combination of all the measurements,
($\underline{u}, \overline{u}$) are the 
actuator bounds, and
($u_{sp}, r_{sp}$) are the input and 
controlled measurement setpoints.
The equality constraint $r_{sp} = H(Cx_s + C_d\dhat)$
can be difficult to satisfy exactly in real time 
depending on the controlled measurement setpoint 
and disturbance estimate. 
Then, the hard setpoint equality constraint 
can be relaxed and moved to the stage cost such 
that the target selector computes a steady state
to minimize the offset in the controlled measurements. 
We assume for the simulation studies
in this article that the input setpoint ($u_{sp}$)
is fixed at some chosen steady state and 
only the controlled measurement setpoint ($r_{sp}$) 
changes in real time.

\subsection{Regulator}
Based on the state estimate and
target steady-state pair,
the following QP is solved to determine the 
actuator move to be applied to the plant
\begin{align} \label{eqn:regulator}
    V_N(\tilde{x}(0), \mathbf{\tilde{u}}) = \dfrac{1}{2}\sum_{k=0}^{N-1} & \Big(|\tilde{x}(k)|^2_Q  + |\tilde{u}(k)|^2_R \Big) + \dfrac{1}{2}|\tilde{x}(N)|^2_P
\end{align}
subject to 
\begin{align}
    \tilde{x}^+ &= A\tilde{x} + B\tilde{u}, \quad \tilde{x}(0) = \xhat - x_s \\
    &\underline{u} \leq \tilde{u} + u_s \leq \overline{u}
    \label{eqn:regulator_constraints}
\end{align}
in which $\xhat$ is the state estimate after the current measurement,
$N$ is the control 
horizon length, and
($\tilde{x}, \tilde{u}$) are the state and control 
in deviation from the current steady-state targets ($x_s, u_s$). 
The penalty matrix $P$ is chosen as the 
optimal cost-to-go matrix of the unconstrained infinite horizon 
linear quadratic problem.
The decision variable in this 
QP is the control sequence 
$\mathbf{\tilde{u}} = [\tilde{u}(0)', \tilde{u}(1)', ..., \tilde{u}(N-1)']'$.
The first move of the optimal solution is denoted 
as the MPC feedback law,
$\kappa_N(\xhat, x_s, u_s) = 
\mathbf{\tilde{u}}^0(0;\xhat, x_s, u_s) + u_s$, 
which is applied to the plant. This feedback law is a function of the 
state estimate and target steady-state pair. A rate-of-change 
penalty on the control can be imposed in the QP objective
by augmenting the dynamic model with a surrogate state 
for the control as discussed in \cite*{rao:rawlings:1999}. 
In this case, 
the control injected at the previous timestep
($u_{-1}$) is supplied to the QP, 
and the feedback law becomes a function of the previous control 
as well in addition to the state estimate
and target steady-state pair.

@@

As reviewed in \cite*{mayne:rawlings:rao:scokaert:2000},
the nominal stability of MPC can be established
using the optimal cost function ($V_N^0(x)$) of the QP
as a Lyapunov function.
The region of attraction of the closed-loop system 
$x^+ = Ax + B\kappa_N(x)$ when no hard terminal region 
constraint is included in the QP
can be characterized as $\mathcal{X}_N = \textnormal{lev}_{Nd + \tau} \
V_N^0(x)$ \citep*{limon:alamo:salas:camacho:2006}.
The parameter $\tau$ is the level set
parameter of the terminal region chosen as
$\bbX_f = \textnormal{lev}_{\tau} \ x'Px$, and $d > 0$
is a constant such that $\ell(x, u) \geq d$ for all 
$x \in \bbR^n \setminus \bbX_f$ and 
$\underline{u} \leq u+u_s \leq \overline{u}$.
The parameter $\tau$ is chosen small enough such that
the unconstrained LQR satisfies the actuator constraints 
for all $x \in \bbX_f$. The optimal cost 
function is continuous and satisfies 
$c_1\norm{x}^2 \leq V_N^0(x) \leq c_2\norm{x}^2$
and $V_N^0(x^+) - V_N^0(x) \leq -c_1\norm{x}^2$ 
for some $c_2 \geq c_1 > 0$.

Assume that a state estimate ($\xhat$) is
used by the NN to compute the control. 
To analyze robustness, the following perturbed
linear system is considered
\begin{align} \label{eq:perturbed_sys}
    \xhat^+ = A\xhat + B\kappa_N(\xhat) + Be_{NN}(\xhat) + w - Ae + e^+
\end{align}
in which $\xhat = x + e$ is the state estimate, 
$w$ is a process disturbance, 
and $e$, $e^+$ are the state estimation errors at the 
current and next timestep respectively. 
We use $\phi_d(k;\xhat)$ to denote a solution of the perturbed system  at time
$k$ starting from an initial state $\xhat$; and $\phi(k;x)$ to denote a
solution  of the nominal system starting  from an initial state $x$ when the
state estimation errors and process disturbances are zero.

\begin{thm} \label{thm:nnrobustness}
For all $ 0 < \rho \leq Nd + \tau$, 
there exist constants $\delta_1, \delta_2, \delta_3 > 0$, 
functions $\beta(\cdot) \in \mathcal{K} \mathcal{L}$
and $\alpha_e(\cdot) , \alpha_w(\cdot), \alpha_{n}(\cdot) \in \mathcal{K}$,
such that for all disturbance sequences
satisfying $\norm{\norm{\mathbf{e}_{k+1}}} \leq \delta_1$,
$\norm{\norm{\mathbf{w}_k}} \leq \delta_2$,
and $\norm{e_{NN}(\xhat)} \leq \delta_3$,
for all $\xhat$ in the set $S := \textnormal{lev}_{\rho} \ V_N^0(\xhat)$,
we have the following bound 
$\norm{\phi_d(k; \xhat)} \leq \beta(\norm{\xhat}, k) 
+ \alpha_e(\norm{\norm{\mathbf{e}_{k+1}}}) + 
\alpha_w(\norm{\norm{\mathbf{w}_k}}) + 
\alpha_n(\bar{e}_{NN})$, 
in which 
$\bar{e}_{NN} = \textnormal{max}_{\xhat \in S} \norm{e_{NN}(\xhat)}$.
\end{thm}

The proof of this theorem along with the required 
input-to-state stability definitions are provided 
in Appendix \ref{app:theorem1}.
In the nominal case, 
the disturbance sequences $\mathbf{e}$ and $\mathbf{w}$ are zero.
The closed-loop states in this case have the bound
$\norm{\phi(k; x)} \leq \beta(\norm{x}, k) + 
\alpha_n(\bar{e}_{NN})$. Any additional value of 
the NN training is to reduce the worst 
approximation error ($\bar{e}_{NN}$) 
in the state space of interest and 
reduce the bound on the closed-loop states.

@@

\begin{itemize}
    \item Controller performance index
    \begin{align*}        
    \Lambda_k = \dfrac{1}{k}\sum_{t=1}^{k} & \big(\norm{x(t) - x_{s} (t)}^2_{Q} \\
        & + \norm{u(t) - u_{s} (t)}^2_{R} + \norm{\Delta u(t)}^2_{S}\big)         
    \end{align*}
    \item \% Performance loss = $100(\Lambda^{\textnormal{F}}_{N_{t}} - \Lambda^{\textnormal{MPC}}_{N_{t}})/\Lambda^{\textnormal{MPC}}_{N_{t}}$
    \item Average and worst-case speedups
    \item Data generation and NN training times
    \item Memory required to store the weights of the NNs
\end{itemize}
Here, $N_t$ is the number of simulation time steps,  
$\Lambda^{\textnormal{F}}_{N_{t}}$ and 
$\Lambda^{\textnormal{MPC}}_{N_{t}}$
are the average stage costs obtained 
at the end of the simulation period by the 
fast controllers and the MPC controller.
The data generation and online timing comparisons are 
performed on a computing cluster that has 
several multi-core CPUs of clock speed 2.4GHz. 
The NN trainings are performed on a 
Tesla V100-SXM2 GPU that has 32 GB memory.

@@

\begin{definition} \label{def:iss}
(Input-to-state stability (ISS)). $\bbW$ is a compact 
set containing the origin and that $X$ is a closed robustly
positive invariant set for the 
system $x^+ = f(x, w)$, $w \in \bbW$. This
system is ISS in $X$ if there exist functions 
$\beta(\cdot) \in \mathcal{K}\mathcal{L}$ and 
$\sigma(\cdot) \in \mathcal{K}$, such that for all
$x \in X$ and $w(i) \in \bbW$, $i \in \bbI_{0:k-1}$, 
$\norm{\psi(k; x, \mathbf{w}_k)} \leq \beta(\norm{x}, k) 
+ \sigma(\norm{\norm{\mathbf{w}_k}})$, in which 
$\psi(k; x, \mathbf{w}_k)$ is the solution of the system 
at time $k$, starting from an initial state $x$, and with the
disturbance sequence $\mathbf{w}_k$ affecting the system.
\end{definition}

\begin{definition} \label{def:iss_lyapunov_func}
(ISS-Lyapunov function). A function $V : X \rightarrow \bbR_{\geq 0}$
is an ISS-Lyapunov function in $X$ for the system $x^+ = f(x, w)$, $w \in \bbW$
if there exist functions $\alpha_1(\cdot), \alpha_2(\cdot), \alpha_3(\cdot) 
\in \mathcal{K}_{\infty}$ and $\sigma(\cdot) \in \mathcal{K}$
such that for all $x \in X$ and $w \in \bbW$
\begin{align*}
    \alpha_1(\norm{x}) \leq & V(x) \leq \alpha_2(\norm{x}) \\
    V(f(x, w)) &- V(x) \leq -\alpha_3(\norm{x}) + \sigma(\norm{w})
\end{align*}
\end{definition}

@@

\begin{prop} \label{prop:continuous_funcs}
    Let $C \subseteq D \subseteq \bbR^n$, $C$ compact, 
    $D$ closed, and $f : D \rightarrow \bbR^n$ continuous.  
    Then there exists a $\sigma(\cdot) \in \mathcal{K}_{\infty}$,
    such that for all $x \in C$ and $y \in D$, we have that 
    $\norm{f(x) - f(y)} \leq \sigma(\norm{x-y})$.
\end{prop}

@@

\textbf{Proof of Theorem \ref{thm:nnrobustness}:}
We proceed in two steps. First, 
we show that there exist constants $\delta_1, \delta_2, \delta_3 > 0$
such that the set $S$ is robustly positive invariant. Then, we 
show that $V_N^0(\xhat)$ is an ISS Lyapunov function for 
the perturbed system \eqref{eq:perturbed_sys} on this set $S$. Denote 
$\tilde{x}^+ = A\xhat + B\kappa_N(\xhat)$ as the nominal 
state evolution using the optimal MPC controller, 
and $d(k) = [ e_{NN}(\xhat(k))'B', w(k)', e(k)', e(k+1)']'$
as the combination of all the disturbances at a given time $k$.

To show robust positive invariance of $S$,  
we establish that if $V_N^0(\xhat) \leq \rho$, 
then $V_N^0(\xhat^+) \leq \rho$.
Since the optimal 
cost function ($V_N^0(\xhat)$) is continuous,
using $\bbR^n$ as the closed set ($\xhat^+ \in \bbR^n$) 
and $S$ as the compact set ($\tilde{x} \in S$),
from proposition \ref{prop:continuous_funcs} we have
for some $\sigma_1(\cdot) \in \mathcal{K}_{\infty}$
\begin{align}
    \norm{V_N^0(\xhat^+) - V_N^0(\tilde{x}^+)} &\leq 
    \sigma_1(\norm{\xhat^+ - \tilde{x}^+})  \nonumber  \\
    &\leq 
   \sigma_1(\norm{B e_{NN}(\xhat)} + \norm{w} 
   \nonumber \\
   & + \norm{A}\norm{e} +  \norm{e^+}) \nonumber \\
     &\leq 
   \sigma_1(\norm{d} + \norm{d} + \norm{A}\norm{d} + \norm{d}) 
   := \sigma_2(\norm{d})
   \label{eq:optimal_cost_bound}
\end{align}
in which $\sigma_2(s) := \sigma_1(\norm{A} s +3s)$, and the last inequality 
follows because $\norm{a} \leq \norm{[a', b']'}$, in which $a$ and $b$
are both vectors, and $\sigma_2(\cdot) \in \mathcal{K}_{\infty}$.
We partition $S$ in outer and inner parts based on 
the value of the optimal cost $V_N^0(\xhat)$, and bound the 
combination of disturbances $d$ in each part 
such that $V_N^0(\xhat^+) \leq \rho$. \\
Case 1: $\rho/2 \leq V_N^0(\xhat) \leq \rho $. Using the 
upper bound on the optimal cost, we have 
$\rho/(2c_2) \leq \norm{\xhat}^2$ for all $\xhat$ 
in this part of $S$. To analyze the worst-case disturbance, 
we have from the equation \eqref{eq:optimal_cost_bound} that 
$V_N^0(\xhat^+) \leq V_N^0(\tilde{x}^+) + \sigma_2(\norm{d}) $.
From nominal MPC cost decrease, we obtain  
$V_N^0(\xhat^+) \leq V_N^0(\xhat) - c_1\norm{\xhat}^2 + 
\sigma_2(\norm{d})$. Thus, if $\sigma_2(\norm{d}) \leq \rho c_1/(2c_2)$, 
at all times in this outer part of $S$, we have 
$V_N^0(\xhat^+) \leq \rho$. \\
Case 2: $V_N^0(\xhat) \leq \rho/2 $. From 
equation \eqref{eq:optimal_cost_bound} and nominal MPC 
cost decrease, we have for this part of $S$ that 
$V_N^0(\xhat^+) \leq \rho/2 + \sigma_2(\norm{d})$.
Thus, if $\sigma_2(\norm{d}) \leq \rho/2$, 
at all times in this inner part of $S$, we have 
$V_N^0(\xhat^+) \leq \rho$.
